### 2015-08-12 Proposals/NERC-CCUS_2015/Outline_ReactiveCCS.tex
Updates in the proposal

\subsection{Background}

@@

The proposed project is focused on predicting \CO\; lateral migration and leakage in complex geological heterogeneity. For the simulation of \CO\; migration it is critically important to accurately represent geological structural characteristics (e.g., facies of the same material property, faults and fractures).  Fully unstructured tetrahedral finite elements have unrivalled geometrical flexibility and are by far the best choice to represent these complex geometries and internal boundaries. Such geological-conforming geometry representation is used with the next generation of models based upon a new family of discontinuous finite element methods. This approach promises a step-change in accuracy from conventional corner-point grids methods in industry-standard flow simulators and represents a novel contribution to CCUS technology. The new model is able to optimally adapt the mesh to represent physics of particular interest, e.g., to focus resolution in the pathways of \CO\; migration.  This model will help quantifying time-scales for supercritical \CO\; migration and potential leakage in geological repositories.

This project aims to study \CO\; plume migration dynamics under reservoir conditions. This and time-dependent supercritical \CO-brine phase change in dedicated geothermal storage sites.
Groundwater flow within overburden strata under dynamic stress conditions

>>>>

%%%
%%% SUBSECTION
%%%
\subsection{Background}

@@

The proposed project is focused on predicting \CO\; plume migration dynamics under reservoir conditions in complex geological heterogeneity. For the simulation of \CO\; migration it is critically important to accurately represent geological structural characteristics (e.g., facies of the same material property, faults and fractures).  Fully unstructured tetrahedral finite elements have unrivalled geometrical flexibility and are by far the best choice to represent these complex geometries and internal boundaries. Such geological-conforming geometry representation is used with the next generation of models based upon a new family of discontinuous finite element methods. This approach promises a step-change in accuracy from conventional corner-point grids methods in industry-standard flow simulators and represents a novel contribution to CCUS technology. The new model is able to optimally adapt the mesh to represent physics of particular interest, e.g., to focus resolution in the pathways of \CO\; migration.  This model will help quantifying time-scales for supercritical \CO\; migration and potential leakage in geological repositories.

%%%
%%% SUBSECTION
%%%
\subsection{Motivation and Objectives}
\CO\; has been injected and stored in underground geological formations to dispose of acidic gasses (mainly H$_{2}$S and \CO) from industrial plants. This has helped to prove that geological \CO\; storage is a viable engineering and economic strategy for reducing GHG emissions. However, geological CO2 storage schemes must comply with requirements (European Council's directive 2009/31/EC) for:
\begin{enumerate}[(a)]
   \item Site screening and selection to identify potential geological locations for \CO\; storage;\label{options:a}
   \item Site characterisation to assess the engineering and economic feasibility;\label{options:b}
   \item Simulation of potential injection strategies to assess (i) pressure and stress distribution and their evolution in time, (ii) plume migration, (iii) oil recovery, (iv) storage capacity and (v) flow rate of injection;\label{options:c}
   \item Risk assessment to identify risks of leakage via pathways identified in (~\ref{options:c}), to design a monitoring scheme for continuous assessment of these risks, and to develop contingency plans in case of  leakage scenarios;\label{options:d}
   \item Site management and operation to ensure that the site is operated in compliance with regulations and to continuously assess the risks associated with the storage site during operation, closure and post-closure stages.\label{options:e} 
\end{enumerate}

Although the main focus of this proposal is to assess potential \CO\; leakage (\ref{options:d}), the technology developed here can also be applied to assess the storage feasibility of new and existing geological sites (\ref{options:b},~\ref{options:c}).  Risk assessment based on detailed numerical simulations can identify specific risk features and processes (e.g., leaky faults and wellbores, unexpected injection pressure rise, residual saturation trapping) which may help prevent unintended CO2 migration. Our model framework may also be used during the continuous site monitoring stage (\ref{options:e}) to assess potential risks during injection and storage as the model can use data from field sensors (data assimilation) to help predict the likelihood of \CO\; leakage.
